Computer-Assisted Analysis of the \\ Anderson-Hájek Ontological Controversy

### 2014-10-27 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Fix minor errors

Although, Hájek did not notice the superfluousness of A4 and A5 in his \AOEH, he did describe yet another emendation (his \AOEHH, cf. Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no claim is made w.r.t. to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive \cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is not only superflous, but also redundant. For A5, no conclusive results were achieved; neither a proof nor a countermodel could be automatically generated. Surprisingly, a countermodel for the weaker A3 was succesfully generated. This is somewhat unsatisfactory (for theistic goals), because it shows that \AOEHH\ does not entail the positiveness of being God-like.

@@

Additionally, Anderson's 14th footnote \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers. Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant). Unfortunately, T3 then becomes unprovable.

The controversy over the superfluousness of A4 and A5 indicates a trend to reduce the ontological argument to its bare essentials. In this regard, 
already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable. These claims have been confirmed by the automated theorem provers. Within the same trend, the alternative proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal}) achieves a high level of minimality. He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being. He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D, A2, A3 and A4. This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K. A4 can be proved in logic \KT (i.e. assuming reflexivity of the accessibility relation). Proving D1 is possible in logic \KFour. Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K. The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB. Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with certain additional presuppositions, to eliminate A5 from his theory as well. Without any additional presupposition, the automated reasoners show that A5 is actually not superfluous. All these results hold for both constant and varying domain semantics.

>>>>

Although, Hájek did not notice the superfluousness of A4 and A5 in his \AOEH, he did describe yet another emendation (his \AOEHH, cf. Appendix \ref{apx:Hajek2}) where A4 and A5 are superfluous (though no claim is made w.r.t. to their redundancy), if A3 is replaced by a stronger axiom (H:A3) additionally stating that the property of actual existence is positive when it comes to God-like beings \cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is not only superflous, but also redundant. For A5, no conclusive results were achieved; neither a proof nor a countermodel could be automatically generated. Surprisingly, a countermodel for the weaker A3 was succesfully generated. This is somewhat unsatisfactory (for theistic goals), because it shows that \AOEHH\ does not entail the positiveness of being God-like.

@@

Additionally, Anderson's 14th footnote \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers. Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant). Unfortunately then, a countermodel for T3 can be found.

The controversy over the superfluousness of A4 and A5 indicates a trend to reduce the ontological argument to its bare essentials. In this regard, 
already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable. These claims have been confirmed by the automated theorem provers (in logic \KFourB). Within the same trend, the alternative proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal}) achieves a high level of minimality. He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being. He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1, A2, A3 and A4. This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K. A4 can be proved in logic \KT (i.e. assuming reflexivity of the accessibility relation). For constant domain semantics, proving D1 is possible in logic \KFour, whereas for varying domain semantics, a counter model can be found even in logic \SFive. Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K. The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB. Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with certain additional presuppositions, to eliminate A5 from his theory as well. Without any additional presupposition, the automated reasoners show that A5 is actually not superfluous. All these results, with the exception of the aforementioned countermodel for D1, hold for both constant and varying domain semantics.
